---
title: "covid 19"
author: "Gloria Jones"
date: "Summer 2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata$sex <- as.factor(kRdata$sex)
kRdata$result <- as.factor(kRdata$result)
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex




Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this

```{r}
dd2 <- kRdata %>% group_by(result,sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=result,y=count,fill=sex))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=result,y=prcnt*100,fill=sex)) 
basicCC + geom_bar(stat="identity", position = "dodge")
````

  
```{r}
barchartGC(~result + sex,data=kRdata)
barchartGC(~result + sex,data=kRdata, type="percent")

```
  
The bar chart indicates that male deaths related to covid make up 60% of the total. Females make up the other 40% of deaths.
  
## Numerical results

```{r}
table1 <- xtabs(~result + sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```
  
The numerical results show out of 100, males make up 60% of covid related deaths, compared to females at 36%. This also shows males are twice as likely to die from covid compared to females.
  
## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The Fisher's exact test for count data shows that males are 2.3 times more likely to die from covdid compared to females. The p-value is also a small number, which allows me to reject the null hypothesis that states covid deaths are independent of sex.
  
## Conclusions

The null hypothesis stated that covid related deaths were independent to sex. First we added a bar chart in order to show the amount of deaths per 100. From there it shows that roughly 60%  deaths out of 100 were males, and females were 40%. Then we ran the numerical results in order to have a more accurate number of covid related deaths between male and female. The results showed that 64 males die to covid compared to 36 females. However, males were more likely to die at a higher rate due to covid compared to females. 

